A Cost-based Approach for Fast Intrusion Detection

### 2017-06-05 tex/paper_v1.tex
change overview

In the approach proposed in this paper, both the supervised learning and ensemble learning methods are employed, namely, the decision tree model and the neural networks model. Empirically, this combination demonstrates a great improvement compared with using single model for prediction.

@@

@@

\begin{figure*}[t]
\begin{centering}
\includegraphics[scale=0.55]{img/overview}
\par\end{centering}
\caption{\label{fig:overview} An overview of the detection framework including the decision tree(top part) and the neural networks(bottom part). The total number of models are calculated on NSL-KDD dataset.}
\end{figure*}

@@

As shown in Fig \ref{fig:overview}, the decision has very simple architecture in the proposed approach.

@@

\subsubsection{Detection Accuracy.}
Tab \ref{tab:detection-accuracy} demonstrates the detection accuracy for each level of features. And apart from the over all accuracy, I take three representative protocol type and service(TCP finger/UDP domain\_u/ICMP eco\_i) as examples.

@@

\subsubsection{False Alarm Rates.}
Tab \ref{tab:detection-false-alarm} demonstrates the false alarm rates for each level of features. As mentioned before, three representative protocol type and service are presented.

@@

\caption{\label{tab:detection-false-alarm}Detection False Alarm Rates.}
\centering{}
\begin{tabular}{ccccc}

@@

The future work for this paper is listed as follows:
\begin{itemize}
\item Calculate the average performance for all the models, not for only some examples.
\item Measure the specific improvement made by distinguishing different level of features quantitatively, such as calculate the exact cost saved for detecting some packets.
\item Try more specific detection results than just normal and anomaly classification. Actually, the NSL-KDD dataset provides the labels of the specific attack types for each connection.
\end{itemize}

\subsection{More Issues}
There are also some practical issues need to be solved for further implementation:
\begin{itemize}
\item There are some protocol type and service with very few training data, such as TCP pm\_dump, UDP tftp\_u and ICMP urh\_i etc. This issue will make it impossible to train a neural network model. Maybe some signature-based technique are needed here.
\item It will cost too much time to train and tune parameters for every model, specifically $72$ models in NSL-KDD dataset. But currently I just train the example models manually.
\item Utilizing the features in NSL-KDD makes the detection in this paper to be connection based, which means that the intrusion can only be detected after all the packages in one connection have arrived. Actually, as suggested in\cite{lee2001real}, some intrusion can be detected only after the first package or a few packages are sent, which is capable to reduce the time cost even more. So extracting more package based features for model training can be considered as another improvement.
\end{itemize}

>>>>

In the approach proposed in this paper, both the supervised learning and ensemble learning methods are employed. The combination of decision tree model and neural networks model demonstrates a great improvement compared with using single model for prediction in NSL-KDD dataset.

@@

\subsubsection{UNB ISCX 2012 Intrusion Detection Evaluation Dataset}
This dataset is proposed in \cite{shiravi2012toward}. According to \cite{shiravi2012toward}, this labeled dataset consists of realistic network and traffic with complete capture. Also, diverse intrusion scenarios are included. This dataset demonstrate a good performance in the comparison with other existing datasets.

@@

\begin{figure*}[t]
\begin{centering}
\includegraphics[scale=0.55]{img/overview-new}
\par\end{centering}
\caption{\label{fig:overview} An overview of the detection framework including the decision tree(top part) and the neural networks(bottom part). The total number of models are calculated on NSL-KDD dataset.}
\end{figure*}

@@

As shown in Fig \ref{fig:overview}, the decision tree has very simple architecture in the proposed approach.

@@

\subsubsection{Detection Accuracy}
Tab \ref{tab:detection-accuracy} demonstrates the detection accuracy for each level of features. And apart from the over all accuracy, I take three representative protocol type and service(TCP finger/UDP domain\_u/ICMP eco\_i) as examples.

@@

\subsubsection{False Alarm Rates}
Tab \ref{tab:detection-false-alarm} demonstrates the false alarm rates for each level of features. As mentioned before, three representative protocol type and service are presented.

@@

\caption{\label{tab:detection-false-alarm}Detection False Alarm Rates}
\centering{}
\begin{tabular}{ccccc}

@@

The future work for this paper is listed as follows:
\begin{itemize}
\item Calculate the average performance for all the models, not for only some examples.
\item Measure the specific improvement made by distinguishing different level of features quantitatively, such as calculate the exact cost saved for detecting some packets.
\item Try more specific detection results than just normal and anomaly classification. Actually, the NSL-KDD dataset provides the labels of the specific attack types for each connection.
\item The experiments are originally planned to conducted on the UNB ISCX IDS 2012 dataset, which is the up-to-date dataset for intrusion detection. However, it may take months to get the dataset following the procedures on the official website. But still, more experiments should be conducted on this new dataset before making any conclusions.
\end{itemize}

\subsection{More Issues}
There are also some practical issues need to be solved for further implementation:
\begin{itemize}
\item There are some protocol type and service with very few training data, such as TCP pm\_dump, UDP tftp\_u and ICMP urh\_i etc. This issue will make it impossible to train a neural network model. Maybe some signature-based technique are needed here.
\item It will cost too much time to train and tune parameters for every model, specifically $72$ models in NSL-KDD dataset. But currently I just train the example models manually.
\item Utilizing the features in NSL-KDD makes the detection in this paper to be connection based, which means that the intrusion can only be detected after all the packages in one connection have arrived. Actually, as suggested in\cite{lee2001real}, some intrusion can be detected only after the first package or a few packages are sent, which is capable to reduce the time cost even more. So extracting more package based features for model training can be considered as another improvement.
\item The datasets that are extensively used currently are still the simulations two decades ago, which makes the models analyzed in these papers unsuitable for detecting newly appeared attacks in recent years. So more datasets are urgently needed in the field of intrusion detection.
\end{itemize}
